Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2015-06-03 tex/parametrized-learning-aoas.tex
clarify text

\cite{Neal:2007zz} also considered the problem of approximating the likelihood function when only a generative model is available. That work sketches a scheme in which one uses a classifier with both $x$ and $\theta$ as an input to serve as a dimensionality reduction map. 
The key distinction comes in the handling of $\theta$.  Neal says  ``we cannot use the classifier on real data, since we don't know the correct value for [$\theta]$'' and goes on to outline an approach where one uses regression on a per-event basis to estimate $\hat{\theta}(x)$ and perform the composition $s(x; \hat{\theta}(x))$ (much like profiling).\footnote{Neal considers a lower-dimensional `bottlenecks' $\theta^*=g(\theta)$, which are not essential to the discussion here.}  This can lead to a significant loss of information since (at least in most particle physics examples) a single event carries little information about the true value of $\theta$, though the full data set $D$ may be informative -- for instance, a single observation would not be sufficient to estimate the variance of a distribution, though repeated observations would.  The work of Neal correctly identifies this as an approximate procedure even in the case of a ideal classifier. In contrast, the approach described here does not eliminate the dependence of the classifier on $\theta$.%
\footnote{As a technical point, in Neal's work, the focus is on approximating the likelihood function (up to a multiplicative constant), which is equivalent to evaluating the ratio  with respect to a fixed $\theta_1$ as in Eq.~\ref{eq:mle_withs}. In Neal's case, the dependence on $\theta$ is eliminated via $s(x; \hat{\theta}(x))$ and the map is constant; however, in this approach the map ratio is explicitly parametrized in terms of $\theta$, so the ratio is important for canceling the corresponding Jacobian factors.} Instead, we embed a parametrized classifier into the likelihood and postpone the evaluation of the classifier to the point of evaluation of the likelihood when $\theta$ is known. This avoid the loss of information that occurs from the regression step $\hat{\theta}(x)$ proposed by Neal and leads to Theorem~1, which is an exact result in the case of an ideal classifier.

>>>>

\cite{Neal:2007zz} also considered the problem of approximating the likelihood function when only a generative model is available. That work sketches a scheme in which one uses a classifier with both $x$ and $\theta$ as an input to serve as a dimensionality reduction map. 
The key distinction comes in the handling of $\theta$.  Neal says  ``we cannot use the classifier on real data, since we don't know the correct value for [$\theta]$'' and goes on to outline an approach where one uses regression on a per-event basis to estimate $\hat{\theta}(x)$ and perform the composition $s(x; \hat{\theta}(x))$ (much like profiling).\footnote{Neal considers a lower-dimensional `bottlenecks' $\theta^*=g(\theta)$, which are not essential to the discussion here.}  This can lead to a significant loss of information since (at least in most particle physics examples) a single event carries little information about the true value of $\theta$, though the full data set $D$ may be informative -- for instance, a single observation would not be sufficient to estimate the variance of a distribution, though repeated observations would.  The work of Neal correctly identifies this as an approximate procedure even in the case of a ideal classifier. In contrast, the approach described here does not eliminate the dependence of the classifier on $\theta$.%
\footnote{As a technical point, in Neal's work, the focus is on approximating the likelihood function (up to a multiplicative constant), which is equivalent to evaluating the ratio  with respect to a fixed $\theta_1$ as in Eq.~\ref{eq:mle_withs}. In Neal's case, the dependence on $\theta$ is eliminated via $s(x; \hat{\theta}(x))$ and the map is constant; however, in this approach the map ratio is explicitly parametrized in terms of $\theta$, so the ratio is important for canceling the corresponding Jacobian factors.} Instead, we embed a parametrized classifier into the likelihood and postpone the evaluation of the classifier to the point of evaluation of the likelihood when $\theta$ is explicitly being tested. This avoids the loss of information that occurs from the regression step $\hat{\theta}(x)$ proposed by Neal and leads to Theorem~1, which is an exact result in the case of an ideal classifier.
